\documentclass[12pt]{amsart}
\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{quiver}
\usepackage{mathrsfs}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{cleveref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=green,      
    urlcolor=cyan,
    citecolor=magenta,
    pdftitle={Perverse Sheaves, Intersection Cohomology, and Mixed Hodge Modules},
    pdfpagemode=FullScreen,
}
\urlstyle{same}


\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\A}{\mathbb{A}}
\newcommand{\sph}{\mathbb{S}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\End}{\text{End}}
\newcommand{\Aut}{\text{Aut}}
\newcommand{\Hom}{\text{Hom}}
\newcommand{\id}{\text{id}}

\DeclareMathOperator{\im}{\text{im}}
\DeclareMathOperator{\coker}{\text{coker}}
\DeclareMathOperator{\D}{\mathcal{D}}


\newcommand{\F}{\mathscr{F}}
\newcommand{\sH}{\mathscr{H}}
\newcommand{\ol}{\overline}
\newcommand{\ul}{\underline}
\newcommand{\wt}{\widetilde}
\newcommand{\pH}{\prescript{p}{}{\mathcal{H}}}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}


\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\numberwithin{equation}{section}




\title{Perverse Sheaves and Intersection Cohomology}
\author{Edward Varvak}

\begin{document}

\maketitle

These are notes I wrote up for my own comprehension while following the text of
Hotta--Takeuchi--Tanisaki \cite{HTT}. No part of this work is original, save for 
any errors I may have introduced throughout my reading.

\section{T-Structures}
\begin{definition}
    Let $\D$ be a triangulated category, and $\D^{\leq 0}, \D^{\geq 0}$ its full subcategories. Set $\D^{\leq n} = \D^{\leq 0}[n]$. The pair $(\D^{\leq 0}, \D^{\geq 0})$ defines a t-structure on $\D$ if
    \begin{itemize}
        \item $\D^{\leq -1} \subseteq \D^{\leq 0}, \D^{\geq 1} \subseteq \D^{\geq 0}$
        \item For all $X \in \D^{\leq 0}$ and $Y \in \D^{\geq 1}$ we have $\Hom_{\D}(X, Y) = 0.$
        \item For any $X \in \D$ there exists an exact triangle
        \begin{equation}
            X_0 \to X \to X_1 \overset{+1}{\to} \cdots
        \end{equation}
        such that $X_0 \in \D^{\leq 0}$ and $X_1 \in \D^{\geq 1}$.
    \end{itemize}
    The full subcategory $\mathcal{C} = \D^{\geq 0} \cap \D^{\leq 0}$ is called the heart of the t-structure $(\D^{\leq 0}, \D^{\geq 0})$.
\end{definition}

The heart of a t-structure allows us to get a good grasp on the ambient derived category.
\begin{theorem}
    $\mathcal{C} = \D^{\geq 0} \cap \D^{\leq 0}$ is an Abelian category, and any exact sequence
    \begin{equation*}
        0 \to X \to Y \to Z \to 0
    \end{equation*}
    gives rise to a distinguished triangle
    \begin{equation*}
        X \to Y \to Z \overset{+1}{\to} \cdots
    \end{equation*}
    in $\D$.
\end{theorem}

This allows us to define the functor $H^0$, which sends our derived category to its heart.
\begin{definition}
    The functor
    \begin{equation*}
        H^0: \D \to \mathcal{C} = \D^{\geq 0} \cap \D^{\leq 0}, \qquad
        X \mapsto \tau^{\leq 0} \tau^{\geq 0}X.
    \end{equation*}
    Likewise, define $H^n(X) = H^0(X[n])$.
\end{definition}
Now let $F: \D_1 \to \D_2$ be a functor of triangulated categories. Denote by $\mathcal{C}_1$, $\mathcal{C}_2$ the hearts of the respective categories.
\begin{definition}
    The additive functor
    \begin{equation*}
        \prescript{p}{}{F}: \mathcal{C}_1 \to \mathcal{C}_2,\qquad
        X \mapsto H^0(F(X)).
    \end{equation*}
    We say that $\prescript{p}{}{F}$ is left t-exact if $F(\D_1^{\geq 0}) \subseteq \D_2^{\geq 0}$, and analogously for right t-exactness.
\end{definition}

\section{Perverse sheaves}

Now we define a t-structure on the derived category of constructable sheaves. Recall that we have a Verdier duality functor $\mathbf{D}_X: D^b_{con}(X)^{op} \to D^b_{con}(X)$.
\begin{definition}
    For $F^\bullet \in \D^b_{con}(X)$, we have $F^\bullet \in \D^{\leq 0}_{con}(X)$ if and only if 
    \[\dim \text{supp}( H^j(F^\bullet)) \leq -j\] for all $j \in \Z$. 
    Likewise, we have $F^\bullet \in \D^{\geq 0}_{con}(X)$ if and only if 
    \[\dim \text{supp}( H^j(\mathbf{D}_X F^\bullet)) \leq -j\] for all $j \in \Z$.
\end{definition}

It is not obvious that this is indeed a t-structure, so it must be stated seperately.
\begin{theorem}
    The pair $(\D_{con}^{\leq 0}, \D_{con}^{\geq 0})$ forms a t-structure. 
\end{theorem}

\begin{definition}
    We call the heart of this t-structure $\text{Perv}(\C_X)$ the perverse sheaves on $X$. The functor $\mathbf{D}_X$ induces a duality of $\text{Perv}(\C_X)$.
    Now, for any functor $F: \D^b_{con}(X) \to \D^b_{con}(Y)$,
    denote by $\prescript{p}{}{F}$ the composition $\prescript{p}{}{H^0} \circ F$, viewed as a functor $\text{Perv}(\C_X) \to \text{Perv}(\C_Y)$.
    In particular, we get the associated functors
    \begin{equation*}
        \prescript{p}{}{f^{-1}},  \prescript{p}{}{f^{!}}: 
        \text{Perv}(\C_Y) \to \text{Perv}(\C_X),
    \end{equation*}
    and
    \begin{equation*}
        \prescript{p}{}{Rf_*}, \prescript{p}{}{Rf_!}:
        \text{Perv}(\C_X) \to \text{Perv}(\C_Y).
    \end{equation*}
\end{definition}



% Suppose that a space $X$ decomposes as $Z \sqcup U$ for a closed $i: Z \to X$ a proper map and $j: U \to X$ an open embedding. Notice that there always exists a natural map
% \begin{equation}
%     i_! \F \to i_* \F,\quad s \mapsto s,
% \end{equation}
% sending a compactly supported section to itself. Clearly, if $i$ is as above, this is an isomorphism for all $\F$, so that $i_! = i_*$.
% Likewise, we have an equality $j^! = j^*$.
% Importantly, we get an exact triangle
% \begin{equation}
%     i_! i^! \to \text{id} \to j_* j^* \overset{[+1]}{\to} \cdots,\qquad 
%     j_! j^! \to \text{id} \to i_* i^* \overset{[+1]}{\to} \cdots.
% \end{equation}

% Now we can define perverse sheaves in terms of this language. Take $X = \bigsqcup_\lambda X_\lambda$ to be a Whitney stratification.
% Denote $i_\lambda: X_\lambda \to X$ the corresponding inclusion, and $d_\lambda$ the dimension of $X_\lambda$.

% \begin{definition}
%     $\sideset{^p}{^{\leq 0}_\lambda}{\D}:= \{
%         \F \in D^b_{const}(X_\lambda)\ |\ \sH^i(\F) = 0\ \forall i > -d_\lambda
%     \}$,\newline
%     $\sideset{^p}{^{\geq 0}_\lambda}{\D}:= \{
%         \F \in D^b_{const}(X_\lambda)\ |\ \sH^i(\F) = 0\ \forall i < -d_\lambda
%     \}$.
% \end{definition}


% With this in mind, we are ready to consider our first example of a perverse sheaf.
% \begin{example}
% Let $X = \mathbb{P}^1$, with $Z = \{p\}$ a point. Then 

% \end{example}

Next, we define minimal extensions of perverse sheaves. Let $X = Z \sqcup U$ with $i: Z \to X$ proper and $j: U \to X$ open.
For $F^\bullet \in \D^b_{con}(X)$, we have natural maps
\begin{equation*}
    j_! F^\bullet \to Rf_* F^\bullet,\qquad s \mapsto s,
\end{equation*}
mapping a section to itself. If $F^\bullet$ is a perverse sheaf on $U$, we also get maps
\begin{equation} \label{eq-minimal-ext}
    \prescript{p}{}{j_!} F^\bullet \to \prescript{p}{}{f_*} F^\bullet.
\end{equation}

\begin{definition}
    We denote by $\prescript{p}{}{f_{!*}} F^\bullet$ the image of \eqref{eq-minimal-ext}, and call it the minimal extension of $F^\bullet \in \text{Perv}(\C_U)$.
\end{definition}
In other words, we have maps
\begin{equation}
    \prescript{p}{}{j_!} F^\bullet \twoheadrightarrow \prescript{p}{}{j_{!*}} F^\bullet
    \hookrightarrow \prescript{p}{}{j_*} F^\bullet.
\end{equation}

\begin{theorem}
    Let $F^\bullet \in \text{Perv}(\C_U)$ as above. Then
    \begin{itemize}
        \item[(i)] $\prescript{p}{}{f_*} F^\bullet$ has no non-trivial sub-objects whose support is contained in $Z$.
        \item[(ii)] $\prescript{p}{}{j_!} F^\bullet$ has no non-trivial quotient object whose support is contained in $Z$.
        \item[(iii)] $\prescript{p}{}{j_{!*}} F^\bullet$ has neither sub- or quotient objects whose support is contained in $Z$.
    \end{itemize}
\end{theorem}

Now suppose that $F^\bullet$ is a simple object in $\text{Perv}(\C_U)$.
We claim that $\prescript{p}{}{j}_{!*} F^\bullet$ is also simple in 
$\text{Perv}(\C_X)$. Indeed, for any subobject 
$G^\bullet \subseteq \prescript{p}{}{j}_{!*} F^\bullet$ we may consideer the exact sequence
\begin{equation}
    0 \to G^\bullet \to \prescript{p}{}{j}_{!*} F^\bullet \to H^\bullet \to 0.
\end{equation}
Since $j^! = j^{-1}$ is exact here, we may apply it to obtain
\begin{equation}
    0 \to j^{-1} G^\bullet \to F^\bullet \to j^{-1} H^\bullet \to 0.
\end{equation}
Since $F^\bullet$ is simple, $j^{-1}G^\bullet$ or $j^{-1} H^\bullet$ is zero,
or in other words, one of $G^\bullet$ and $H^\bullet$ is supported in $Z$.
By the previous theorem, this makes it zero.

\begin{theorem} \label{minimal-characterization}
    $G = \prescript{p}{}{j_{!*}} F^\bullet$ is the unique perverse sheaf 
    satisfying the conditions
    \begin{itemize}
        \item[(i)] $G^\bullet|_U \simeq F^\bullet$,
        \item[(ii)] $i^{-1} G^\bullet \in 
        \prescript{p}{}{\D}^{\leq -1}_{con}(Z)$,
        \item[(iii)] \(i^! G^\bullet \in 
        \prescript{p}{}{\D}^{\geq 1}_{con}(Z)\).
    \end{itemize}
\end{theorem}




\section{Intersection cohomology}

Let $X$ be an irreducible complex projective variety (or an irreducible compact analytic space) of dimension $d$. We may start by defining the intersection cohomology complex in the following way.
\begin{definition}
    Take $U \subseteq X^{\text{reg}}$ to be a Zariski open subset of the smooth locus of $X$. For $\underline{\C}_U[d]$ the constant perverse sheaf on $U$, define $IC_X^{\bullet}$ to be a minimal extension to all of $X$.
\end{definition}

We may explicitly construct $IC_X^{\bullet}$. Fix a Whitney stratification $X = \bigsqcup_\alpha X_{\alpha}$, 
\begin{equation}
    X_k = \coprod_{\dim X_\alpha \leq k} X_\alpha, \quad U_k := X \smallsetminus X_{k-1},
\end{equation}
with each $X_k \smallsetminus X_{k-1}$ a smooth $k$-dimensional complex manifold. We have a family of maps $j_k: U_k \hookrightarrow U_{k-1}$, for $k=1,\dots, d$. Altogether, this fits into
\begin{equation}
    \emptyset \hookrightarrow U_d \hookrightarrow U_{d-1} \hookrightarrow \cdots
    \hookrightarrow U_1 \hookrightarrow X.
\end{equation}

\begin{theorem} \label{ic-structure}
    \(IC_X^{\bullet}\) is quasi-isomorphic to the complex
    \begin{equation}
        \prescript{p}{}{j_{!*}}(\C_U[d_X]) \simeq 
        \left(\tau^{\leq -1}Rj_{1*}\right) \circ \left(\tau^{\leq -2}Rj_{2*}\right)
        \circ \cdots \circ \left(\tau^{\leq -d}Rj_{d*}\right)
        \left(\underline{\C}_U[d]\right).
    \end{equation}
\end{theorem}
\begin{proof}
Suppose that $F^\bullet$ is a perverse sheaf on $U_k$ whose restriction to 
any strata $X_\alpha \subseteq U_k$ has locally constant $\sH^i$.
We show that $G^\bullet := \tau^{\leq -k} Rj_{k*}(F^\bullet)$ satisfies the
unique characterization of a minimal extension from Theorem 
\Cref{minimal-characterization}. First, note that $U_k$ consists of 
strata of dimension $\geq k$, so that \(\sH^r(F^\bullet)  = 0\) for 
\(r > -k\). In particular,
\begin{equation*}
    \left[\tau^{\leq -k} Rj_{k*} F^\bullet\right]|_{U_k} \simeq F^\bullet,
\end{equation*}
so that what we have is indeed an extension of $F^\bullet$, satisfying (i).
Next, set \(Z := U_{k-1} \smallsetminus U_k = 
\bigsqcup_{\dim X_\alpha = k-1} X_\alpha\). Denote by $i: Z \to U_k$ the
associated closde embedding. Then $i^{-1} G^\bullet$ has locally constant 
cohomology sheaves on each $X_\alpha \subseteq Z$, so that 
\(\sH^r(i^{-1} G^\bullet) = 0\) for \(r > -k\). This implies that 
\(i^{-1} G^\bullet \in \prescript{p}{}{\D}_{con}^{\leq -1},\) thus satisfying
condition (ii). Finally, consider the triangle
\begin{equation}
    G \to Rj_{k*} F^\bullet \to \tau^{\geq -k+1} Rj_{k*} F^\bullet
    \overset{+1}{\to} \cdots
\end{equation}
Applying $i^!$, we note that the middle vanishes. This gives us isomorphisms
\begin{equation*}
    i^! G^\bullet \simeq i^! (\tau^{\geq -k+1} Rj_{k*} F^\bullet)[-1]
\end{equation*}
so that $\sH^r(i^! G^\bullet) = 0$ for \(r < -k\). But since \(i^! G\) has locally
constant cohomology on \(X_\alpha \subseteq Z\), we get
\(i^! G \in \prescript{p}{}{\D}_{con}^{\geq 1}(Z)\), proving (iii).
\end{proof}

Importantly, the intersection cohomology complex is self-dual in the way we expect of perverse sheaves.

\begin{theorem}\label{canonical-morphisms}
    \(IC_X^{\bullet} \simeq \mathbb{D}_X (IC_X^{\bullet}) \). Furthermore, there exist canonical morphisms
    \begin{equation}
        \underline{\C}_X \to IC_X^{\bullet}[-d] \to \omega_X^\bullet[-2d].
    \end{equation}
\end{theorem}
\begin{proof}
We have an isomorphism in the derived category of constructible sheaves
from above
\begin{equation}
    \tau^{\leq -d}\ \prescript{p}{}{j}_{!*} \underline{\C}_X[d] \simeq
     Rj_{1*} \circ \cdots \circ Rj_{d*}
     \left(\underline{\C}_X[d]\right) \simeq (j_* \C)[d].
\end{equation}
This is none other than the intersection complex of $X$, and we see that it admits a canonical map from $\underline{\C}_X$. The second map in the composition comes from taking the Verdier dual,
\(IC_X^\bullet[d] \to \omega_X^\bullet \).
\end{proof}

With the intersection complex in place, we are ready to define intersection cohomology.
\begin{definition}
    For $i \in \Z$, we define
    \begin{equation}
        IH^i(X) = H^i\left(R\Gamma\left(X, IC_X^\bullet[-d]\right)\right),
    \end{equation}
    \begin{equation}
        IH_c^i(X) = H^i\left(R\Gamma_c\left(X, IC_X^\bullet[-d]\right)\right).
    \end{equation}
\end{definition}

Importantly, the intersection cohomology of $X$ satisfies Poincare duality.
\begin{theorem}
    Let $X$ be irreducible of dimension $d$. Then
    \begin{equation}
        IH^i(X) \simeq \left[IH_c^{2d-i}(X)\right]^*.
    \end{equation}
\end{theorem}
\begin{proof}
Let $a_X: X \to \{p\}$ be the unique map to a point. We have 
\begin{equation}
    R\Hom_\C(Ra_{X!} IC^\bullet, \C) \simeq 
    Ra_{X*} R\Hom_{\C_X}(IC_X^\bullet, \omega_X^\bullet).
\end{equation}
by Verdier duality. But \(IC_X^\bullet\) is self-dual, so that
\begin{equation*}
    R\Hom_{\C_X}(IC_X^\bullet, \omega_X^\bullet) = 
    \mathbf{D}_X(IC_X^\bullet) = IC_X^\bullet.
\end{equation*}
Thus we get an isomorphism
\begin{equation}
    \left[R \Gamma_c(X, IC_X^\bullet)\right]^* \simeq 
    R\Gamma(X, IC_X^\bullet).
\end{equation}


\end{proof}

\begin{example}
Let $X$ be an irreducible complex projective variety with isolated singular points $p_1,\dots, p_k$. Then it suffices to consider the stratification
\[ X = \{p_1,\dots,p_k\} \sqcup X^{\text{reg}}. \]
Then we have \(X_0 = \dots = X_{d-1} = \{p_1,\dots, p_k\}\), and \(X_d = X\). Thus, we get the inclusions of complex manifolds
\[ \emptyset \to U_d = X^{\text{reg}} = U_1 \overset{j_1}{\to} X.\]
The only interesting map in this stratification is $j_1$, the inclusion of the smooth locus. In particular, by Theorem \ref{ic-structure},
\[ IC_X^\bullet \simeq \tau^{\leq -1}\left(Rj_{1*} \underline{\C}_{U_1}\right). \] 
Now, associated to truncation is the exact couple
\begin{equation}
    IC_X^\bullet[-d] \to Rj_{1*} \underline{\C}_{U_1} \to 
    \tau^{\geq d}\left(Rj_{1*} \underline{\C}_{U_1}\right) \overset{+1}{\to} \cdots
\end{equation}
But now we can apply $R\Gamma$ and take cohomology. For $0 \leq i < d$, we have
\( IH^i(X) = H^i(X, Rj_{1*}\overline{\C}_{U_1}) = H^i(X^{\text{reg}}, \C). \) For $i=d$, we do not get a clear vanishing on the right, but the map
\(IH^d(X) \to H^d(X^{\text{reg}}, \C)\) is injective. To analyze this case more carefully, consider the canonical morphism
\( \underline{\C}_X \to IC_X^\bullet[-d] \)
guarenteed by Theorem \Cref{canonical-morphisms}.

We may associate to this natural map a new exact couple
\begin{equation}
    \underline{\C}_X \to IC_X^\bullet[-d] \to F^\bullet \overset{+1}{\to} \cdots
\end{equation}
Here \(F^\bullet\) is a constructable sheaf supported only on the zero-dimensional closed subset $\{p_1, \dots, p_k\} = X_0$. In particular, \(H^i(F^\bullet) = 0\) for all $i \geq d$. Thus, we get \(IH^i(X) = H^i(X, \C)\) for $d < i \leq 2d$. After $i = 2d$, the cohomology of $\underline{\C}_X$ vanishes, and with it $IH^i(X)$. Summarizing,
\begin{equation}
    IH^i(X) = \begin{cases}
        H^i(X^{\text{reg}}, \C) & : 0 \leq i < d \\
        \text{im}(H^i(X, \C) \to H^i(X^{\text{reg}}, \C)) & : i = d \\
        H^i(X, \C) & : d < i \leq 2d \\
        0 & : \text{otherwise}
    \end{cases}
\end{equation}
\qed
\end{example}


\section{Computations via Saito's mixed Hodge modules}
To compute intersection complexes of standard perverse sheaves, we use Saito's 
mixed Hodge modules (\cite{saito88-HM}, \cite{saito90-MHM}). For a quick summary of 
this theory, refer to \cite[\S 2.1]{deC-Rap-Sac21}. Throughout these computations,
we loosely follow \cite[\S 2]{deC-Mig09}.

Let \(X\) be a variety over \(\C\). Saito \cite{saito90-MHM} defines a category
\(\D^b \mathrm{MHM}(X)\) of complexes of mixed Hodge modules, equipped with
Hodge-theoretic weight formalism and a natural morphism
\(\mathrm{rat}: \D^b \mathrm{MHM}(X) \to \D^b_{con}(X, \Q)\). This category 
has two t-structures: the one corresponding to the standard t-structure on 
\(\D^b_{con}(X, \Q)\), and the perverse t-structure. This sets a heart:
if \(K \in \mathrm{MHM}(X)\), then \(\mathrm{rat}(K) \in \mathrm{Perv}(X, \Q)\).
When restricted to a suitable dense open set, the objects in \(\mathrm{MHM}(X)\) 
become admissable variations of polarizable mixed Hodge structures.

More importantly, for \(f: X \to B\) a proper morphism of varieties and 
\(K \in MHM(X)\), if the derived pushforwards (with respect to the standard t-structures)
\(R^i f_* K\), \(R^i f_! K\) are perverse sheaves, they may be endowed with the 
structure of an object in \(\D^b \mathrm{MHM}(X)\). More concretely, 
we have the following theorem.

\begin{theorem}[Decomposition Theorem]
    Let \(K \in \mathrm{MHM}(X)\) be of pure weight \(w\). Then there is a splitting 
    of \(Rf_* K\) with respect to the perverse t-structure 
    \begin{equation}
        Rf_* K = \bigoplus_{k} \pH^k(Rf_* K)[-k],
    \end{equation}
    where each summand \(\pH^k(Rf_* K)[-k]\) is pure and semisimple in 
    \(\D^b \mathrm{MHM}(X)\) of weight \(w-k\). (Should there be no \([-k]\) shift?) 
    Moreover, if \(\eta\) is an ample class on \(X\), there is a relative Hard 
    Lefschetz isomorphism
    \begin{equation}
        \eta^i: \pH^{-i}(Rf_* K) \simeq \pH^i(Rf_* K).
    \end{equation}
\end{theorem}

\subsection{Fibrations of smooth algebraic varieties.}
In this section, we focus on the case of \(f: X \to B\), where \(f\) is proper with 
positive-dimensional fibers, and \(X\) and \(B\) are smooth. 

\begin{example}
Let \(X\) be a projective surface and \(B\) a curve. Denote by \(B^\circ\) the locus 
of regular values of \(f\), \(\Delta = B \smallsetminus B^\circ\). Then
\(R^1 f_* \Q_X|_{B^\circ}\) is a weight one variation of Hodge structures. Denote 
the underlying local system by \(\mathbb{V}\).

In order for \(\mathbb{V}\) to appear in middle perversity in the decomposition theorem, 
we must shift the pushforward by \(2\): set \(W = Rf_* \Q_X[2]\).
\begin{align*}
    W|_{B^\circ} &= \pH^{-2}(W|_{B^\circ})[2] \oplus 
    \pH^{-1}(W|_{B^\circ})[1] \oplus \pH^0(W|_{B^\circ})[0] \\
    &= \Q_{B^\circ}[2] \oplus \mathbb{V}[1] \oplus \Q_{B^\circ}.
\end{align*}

Next, we set our attention to the singular fibers of the pushforward. Let 
\(p \in \Delta\) be a critical value. Then
\begin{align*}
    \pH^{-2}(W)_p[2] &= H^0(X_p, \Q)_p[2] = \Q_p[2], \\
    \pH^{-1}(W)_p[1] &= H^1(X_p, \Q)_p[1], \\
    \pH^{-2}(W)_p[0] &= H^2(X_p, \Q)_p[0] = \bigoplus_{X_p^{(i)} \subseteq X_p}
    \left<[X_p^{(i)}]\right>_p[0],
\end{align*}
where the last sum is taken over the irreducible components of \(X_p\).

Altogether, the decomposition theorem yields
\begin{equation} \label{eq-example-decomp-wt1-loc}
    W = \Q_B[2] \oplus \ol{\mathbb{V}}[1] \oplus \Q_B[0] \oplus \bigoplus_{p \in \Delta} 
    \left(\bigoplus_{X_p^{(i)} \subseteq X_p} 
    \frac{\left<[X_p^{(i)}]\right>}{\left<[X_p]\right>}\right)_p[0].
\end{equation}
Here \(\ol{\mathbb{V}} = R^1 f_* \Q_X\). We may also compute the intersection complex 
\(IC_B(\mathbb{V})\) valued in the local system \(\mathbb{V}\). It appears in this 
decomposition, and indeed, it is clear from the perverse degrees that 
\begin{equation}
    IC_B(\mathbb{V}) = \ol{\mathbb{V}}[1] \oplus \bigoplus_{p \in \Delta} 
    \left(\bigoplus_{X_p^{(i)} \subseteq X_p} 
    \frac{\left<[X_p^{(i)}]\right>_p}{\left<[X_p]\right>_p}\right)[0].
\end{equation}
This is identified with \(R^1 f_* \Q_X[1]\) if and only if the fibers of \(f\) are 
all irreducible.
\qed
\end{example}


\begin{example}
We now consider the same example, but on the level of mixed Hodge modules. As a Hodge
module, \(\Q_X(0)\) is pure of weight 0. Over \(B^\circ\), its pushforward 
decomposes as 
\begin{equation}
    \Q_{B^\circ}(0)[0] \oplus \mathcal{V}[-1] \oplus \Q_{B^\circ}(-1)[-2],
\end{equation}
where \(\mathcal{V}\) is the weight one variation of Hodge structures associated to the 
local system \(\mathbb{V}\) above. Over \(p \in \Delta\), we get 
\begin{equation}
    Rf_* \Q_X|_p = H^0(X_p, \Q)[0] \oplus H^1(X_p, \Q)[-1] \oplus H^2(X_p, \Q)[-2].
\end{equation}
Hence, we get the same decomposition as in \eqref{eq-example-decomp-wt1-loc}: 
\begin{equation}
   Rf_* \Q_X = \Q_B(0)[0] \oplus \ol{\mathcal{V}}[-1] \oplus \Q_B(-1)[-2] \oplus
   \bigoplus_{p \in \Delta} \left(\frac{\left<[X_p^{(1)}], \dots, [X_p^{(n)}]\right>}
   {\left<[X_p^{(1)} + \cdots + X_p^{(n)}]\right>}\right)_p[-2].
\end{equation}
Notice that the relative Hard Lefschetz theorem sends \(\Q_B(0)\) to \(\Q_B(-1)[-2]\)
isomorphically, and fixes all other primative cohomology. We get a mixed Hodge module 
associated to the intersection complex, 
\begin{equation}
    \mathscr{IC}_B(\mathcal{V}) = \ol{\mathcal{V}}[-1] \oplus 
    \bigoplus_{p \in \Delta} \left(\frac{\left<[X_p^{(1)}], \dots, [X_p^{(n)}]\right>}
   {\left<[X_p^{(1)} + \cdots + X_p^{(n)}]\right>}\right)_p[-2].
\end{equation}
Here \(\ol{\mathcal{V}} = R^1 f_* \Q_X\). From this description, we see that it is a
simple weight one mixed Hodge module if and only if the fibers of \(f\) are irreducible.

We may compute \(IH^k(B, \mathcal{V}) = H^k(R \Gamma(\mathscr{IC}_B(\mathcal{V})))\).
By Saito, these carry mixed Hodge structures. In this case, we have 
\(IH^1(B, \mathcal{V}) = H^0(\ol{\mathcal{V}})\) the non-vanishing invariant cycles, and
\(IH^2(B, \mathcal{V}) = \bigoplus_{p \in \Delta} H^2(X_p, \Q)^{\mathrm{prim}}\)
the primative weight two cohomogy of the singular fibers.
\qed
\end{example}




\subsection{Resolutions of normal isolated singularities}
We next consider another important class of morphisms: Let \(X\) be a projective 
variety singular at \(x\), and let \(f: \wt{X} \to X\) be a resolution. We will 
use \(f\) compute \(IC_X\). In particular, we may recover \(IH^*(X, \Q)\), which 
satisfies Poincare duality.

\begin{example}
Let \(C \subseteq \mathbb{P}^2\) be a plane curve, and consider \(X\) the cone
over \(C\). We have a natural resolution \(f: \wt{X} \to X\) given by the blow-up 
morphism, \(\wt{X} = \mathrm{Bl}_x X\). Consider \(K = Rf_* \Q_{\wt{X}}\).

\(\Q_{\wt X}(0)\) is a weight zero Hodge module on \(\wt{X}\), so its pushforward 
\(K \in \D^b\mathrm{MHM}(X)\) is a complex of pure weight zero. The restriction 
to the smooth locus of \(X\) is \(K|_{X^\circ} = \Q_{X^\circ}(0)\), a local system.
Over the singular point, the fiber is \(\wt{X}_x \simeq S\), so its cohomology is
\begin{equation}
    K_x = \Q_x(0)[0] \oplus H^1(C, \Q)_x[-1] \oplus \Q_x(-1)[-2].
\end{equation}
Thus, the decomposition theorem yields 
\begin{equation}
    K = \Q_X(0)[0] \oplus H^1(C, \Q)_x[-1] \oplus \Q_x(-1)[-2].
\end{equation}

Now, the first summand is supported in dimension one, while the second is supported 
in dimension zero. We therefore see that both of them of the same perverse degree,
so that 
\begin{equation}
    \mathscr{IC}_X = \Q_X(0)[0] \oplus H^1(C, \Q)_x[-1].
\end{equation}
In particular, this implies that 
\begin{equation}
    IH^i(X, \Q) = \begin{cases}
        H^1(C, \Q) & : i = 1 \\
        H^i(X, \Q) & : i \neq 1
    \end{cases}
\end{equation}
Intersection cohomology satisfies Poincare duality, so we can mostly recover 
\(H^*(X, \Q)\) from this description:
\begin{equation}
    H^0(X, \Q) = \Q,\ H^1(X, \Q) = 0,\ H^2(X, \Q) = ?,\ H^3(X, \Q) = \Q^{2g},\ 
    H^4(X, \Q) = \Q.
\end{equation}
We can check that \(H^2(X, \Q) = \Q\) by noting that all non-ruling classes in 
\(H^2(X, \Q)\) can be contracted to zero at the cone point, and are therefore trivial.
\qed
\end{example}


\begin{example}
This time, set \(S \subseteq \mathbb{P}^2\) to be a quadric surface, and consider
the cone \(X \subseteq \mathbb{P}^3\) over \(S\). We have two resolutions to 
choose from: the blow-up of the cone point \(x \in X\), or the blow-up of a line 
\(L \ni x\) in \(X\).

First, let us consider the former: \(\wt{X} = \mathrm{Bl}_x X\), \(f: \wt{X} \to X\). 
As before, let \(K = Rf_* \Q_X\). We once again have \(K|_{X^\circ} = \Q_{X^\circ}(0)\).
Anything interesting will again have to happen over the cone point. The fiber 
\(\wt{X}_x \simeq S\), which happens to be simply connected. We see that 
\begin{equation}
    K = \Q_X(0)[0] \oplus H^2(S, \Q)_x[-2] \oplus \Q_x(-2)[-4].
\end{equation}
The first summand is supported in dimension two, while the primitive part of the second 
is supported in dimension zero. Thus, they have the same perversity, and 
\begin{equation}
    \mathscr{IC}_X = \Q_X(0)[0] \oplus H^2(S, \Q)_x^{\mathrm{prim}}[-2].
\end{equation}

Now we can again compute the intersection cohomology of \(X\). We get 
\begin{equation}
    IH^i(X, \Q) = \begin{cases}
        H^2(X, \Q) \oplus \left<L_1 - L_2\right> & : i = 2\\
        H^i(X, \Q) & : i \neq 2.
    \end{cases}
\end{equation}
In particular, we can conclude that 
\begin{equation}
    H^0(X, \Q) = \Q,\ H^2(X, \Q) = \Q,\ H^4(X, \Q) = \Q^2,\ H^6(X, \Q) = \Q, 
\end{equation}
and all odd-degree cohomology vanishes from standard rationality arguments.
\qed
\end{example}

\begin{example}
Had we instead taken \(\wt{X} = \mathrm{Bl}_L X\), \(f: \wt{X} \to X\), we would have 
\(\wt{X}_x = \mathbb{P}^1\). Now if we consider \(K = Rf_* \Q_{\wt{X}}\), the 
decomposition theorem yields 
\begin{equation}
    K = \Q_X(0)[0] \oplus \Q_x(-1)[-2] \simeq \mathscr{IC}_X.
\end{equation}
Hence, we see that some resolutions are better for computing intersection cohomology 
than others. Note that this map \(f\) had the property that 
\begin{equation}
    Rf_* \mathscr{IC}_{\wt{X}} = \mathscr{IC}_X.
\end{equation}
\qed
\end{example}





\bibliographystyle{alpha}
\bibliography{refs}


\end{document}
